`r params$doc_title`

### 2022-11-23 scripts/MSc.Rmd
added MC4R association

filter_FTO <- rsIds_GeneNames %>% 
  filter(str_detect(associated_gene, pattern = "FTO")) %>% 
  distinct(refsnp_id, .keep_all = TRUE)

t(filter_FTO) %>% kable() %>% 
  kable_styling(font_size=10)
```

@@

my_data_harm <- as_tibble(cbind(my_data_harm, pvals_df))

# 5. Calculate cumulative explained variances
# 6. Represent figure similar to Fig1 in Chatterjee’s paper
```

@@

```{r calculate-explained-variance}
# outsourced to "./scripts/PVE_calculation_fns.R"
```

@@

saveRDS(est, paste0("../output/Rdata/", Sys.Date(), "_est.rds"))

@@

```{r simulatesimulatesimulate, warning=FALSE}

# data FOR simulation in simulation script
saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
sim_vars <- list(n=n, colnames_pval=colnames_pval)
saveRDS(sim_vars, paste0("../output/Rdata/", Sys.Date(), "_sim_vars.rds"))

@@

# read dat from "simstudy.R"
est_sim <- readRDS(file = "../output/Rdata/2022-11-18_est_sim.rds")

est_sim <- readRDS(file = "../output/Rdata/2022-11-22_est_sim_NSIM_5.rds")
```

@@

```{r plot-performance-metrics}
#BIAS
ggplot(performance_metrics, aes(x=N, y=bias))+ 
  geom_point()+
  scale_x_continuous(labels = comma)+
  theme_few()

#MSE
ggplot(performance_metrics, aes(x=N, y=mse))+ 
  geom_point()+
  scale_x_continuous(labels = comma)+
  theme_few()

# compare distribution of simulated thetas against theta_324000
boxplot_dat <- est_sim %>% 
  select(!!!(colname_theta)) %>% 
  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
  mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))

ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
  geom_boxplot()+
  geom_jitter(width=0.1,alpha=0.2)+
  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
  annotate(geom = "text", x = "300000", y = args$theta+0.5,
           label= expression( paste("ref: ", theta[X] ))  , 
           color= ggthemes_data$few$colors$Dark[9,2][[1]]
             )+
  scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
  ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
  labs(title = "Estimates from beta simulation",
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       )+
  coord_flip()+
  theme_few()
```

```{r}
# EXPOSURE PVE
est_sim %>% 
  select(!!!(colname_pve_X)) %>% 
  pivot_longer(everything(), names_to = "PVE_X", values_to = "value_X") %>% 
  cbind(sample_size_filtered) %>% as_tibble() -> pve_X_sim

@@

fig01b.update +
  geom_boxplot(data=boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, group=sample_size_filtered))

  geom_boxplot(data = test, aes(x=sample_size_filtered, y=value*100, # *100 to get percent values
                                group=sample_size_filtered), 
               col=ggthemes_data$few$colors$Dark[3,2][[1]],
               alpha=0.6)+

ggplot(test, aes(x=as.factor(sample_size_filtered), y=value))+
  geom_boxplot()

plot_data1b <- df_cum_expl_var %>% 
  pivot_longer(cols = c(cum_expl_varX2, cum_expl_varU, cum_expl_varY)) %>% 
  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")),
         name = factor(name, levels = c("% var(X)", "% var(Y)", "% var(U)") )
         )
## plot
ggplot(plot_data1b, aes(x=n, y=value, col=name))+
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                     labels = comma, 
                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
                     limits = c(10000, 300000))+
  #ylab("% variance explained by IVs")+
  labs(title = "Figure 1",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
    limits = c(0,5)
  ) + 
  #geom_boxplot(data=boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, group=sample_size_filtered))
  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, # *100 to get percent values
                                group=sample_size_filtered), 
               col=ggthemes_data$few$colors$Dark[3,2][[1]],
               alpha=0.2, size=0.2)+
  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_Y*100, # *100 to get percent values
                                group=sample_size_filtered), 
               col=ggthemes_data$few$colors$Dark[4,2][[1]],
               alpha=0.2, size=0.2)+
  theme_few()+
  #scale_colour_few() +
  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
  #                                         "% var(X)" = "blue", 
  #                                         "% var(Y)" = "darkblue", 
  #                                         "Number of IVs" = "black"
  #                                         ))
  #
  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                           "Number of IVs" = "black"),
                     # order
                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
  )
```

>>>>

filter_FTO <- rsIds_GeneNames %>% 
  filter(str_detect(associated_gene, pattern = "FTO") | str_detect(associated_gene, pattern = "MC4R")) %>% 
  distinct(refsnp_id, .keep_all = TRUE)

t(filter_FTO) %>% kable() #%>% 
  kable_styling(font_size=10)
```

@@

my_data_harm <- as_tibble(cbind(my_data_harm, pvals_df))
```

@@

```{r calculate-explained-variance}
# outsourced to "./scripts/PVE_calculation_fns.R"

## CALCULATE explained variance
my_data_harm <- my_data_harm %>% 
  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf.bmi, beta = beta.bmi), 
         explained_variance_BMI2 = explained_variance_numeric2(maf=eaf.bmi, 
                                                           beta=beta.bmi, 
                                                           se_beta=se.bmi, 
                                                           samplesize=n[length(n)]

         ),
         explained_variance_GSD = explained_variance_binary(PA = eaf.gsd, 
                                                            RR1 = exp(beta.gsd), 
                                                            RR2 = exp(beta.gsd)^2, 
                                                            K = 0.1  # assumed prevalence in population
                                                            )$Vg,
         explained_variance_GBC = explained_variance_binary(PA = eaf.gbc, 
                                                            RR1 = exp(beta.gbc), 
                                                            RR2 = exp(beta.gbc)^2, 
                                                            K =1.2/100000  # assumed prevalence in population
                                                            )$Vg,
         explained_variance_GBC2 = explained_variance_binary(PA = eaf.gbc, 
                                                            RR1 = exp(beta.gbc), 
                                                            RR2 = 1+(2*(exp(beta.gbc)-1)), 
                                                            K =1.2/100000  # assumed prevalence in population
                                                            )$Vg

         )
```

@@

saveRDS(est, paste0("../output/Rdata/", Sys.Date(), "_est.rds"))
```

@@

```{r simulatesimulatesimulate, warning=FALSE}
# data FOR simulation in simulation script
saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))
sim_vars <- list(n=n, colnames_pval=colnames_pval)
saveRDS(sim_vars, paste0("../output/Rdata/", Sys.Date(), "_sim_vars.rds"))

@@

# read dat from "simstudy.R"  #------------------------------------------------
est_sim <- readRDS(dplyr::last(list.files(path = "../output/Rdata", pattern = "_est_sim_NSIM_", full.names = TRUE)))
```

@@

```{r plot-performance-metrics-bias-mse}
#BIAS
ggplot(performance_metrics, aes(x=N, y=bias))+ 
  geom_point()+
  scale_x_continuous(labels = comma)+
  theme_few() -> biasplot

#MSE
ggplot(performance_metrics, aes(x=N, y=mse))+ 
  geom_point()+
  scale_x_continuous(labels = comma)+
  theme_few() -> mseplot

ggarrange(biasplot, mseplot) ->bias_mse_plot

filename_simplot_bias_mse <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_bias_mse.png")
ggsave(filename = filename_simplot_bias_mse, plot = bias_mse_plot, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

```{r plot-performance-metrics-boxplot}
# compare distribution of simulated thetas against theta_324000
boxplot_dat <- est_sim %>% 
  select(!!!(colname_theta)) %>% 
  pivot_longer(cols = everything(), names_to = "theta_sim") %>% 
  mutate(theta_sim_numeric = format(as.numeric(stringr::str_extract(string =theta_sim, pattern = "\\d+")), scientific=FALSE))

ggplot(boxplot_dat, aes(x=forcats::fct_rev( as.factor(theta_sim_numeric)), y=value)) +
  geom_boxplot()+
  geom_jitter(width=0.1,alpha=0.2)+
  geom_hline(yintercept=  args$theta, col=ggthemes_data$few$colors$Dark[9,2][[1]], alpha=0.6, linetype=2, size=0.8)+
  annotate(geom = "text", x = "300000", y = args$theta+0.5,
           label= expression( paste("ref: ", theta[X] ))  , 
           color= ggthemes_data$few$colors$Dark[9,2][[1]]
             )+
  scale_x_discrete(name = paste0("Size of sample ", as.roman(1)),  breaks = c(" 50000", "100000", "150000", "200000", "250000", "300000"), )+ 
  ylab(expression(paste(theta[sim], " with ", n[sim]== 25 ) )  ) +
  labs(title = "Estimates from beta simulation",
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       )+
  coord_flip()+
  theme_few() -> simplot_betas
simplot_betas

filename_simplot_betas <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_betas.png")
ggsave(filename = filename_simplot_betas, plot = simplot_betas, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

```{r simulated-pves}
# EXPOSURE PVE
est_sim %>% 
  select(!!!(colname_pve_X)) %>% 
  pivot_longer(everything(), names_to = "PVE_X", values_to = "value_X") %>% 
  cbind(sample_size_filtered) %>% as_tibble() -> pve_X_sim

@@

## plot
ggplot(plot_data1b, aes(x=n, y=value, col=name))+    # plot_data1b from section Simlating on empirical data (PVE)
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
  scale_x_continuous(name = paste0("Size of sample ", as.roman(1)),  # 2-sample-MR or label "Study Size"
                     labels = comma, 
                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
                     limits = c(10000, 300000))+
  #ylab("% variance explained by IVs")+
  labs(title = "PVE with simulation added",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs"),
    limits = c(0,10)
  ) + 
  #geom_boxplot(data=boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, group=sample_size_filtered))
  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_X*100, # *100 to get percent values
                                group=sample_size_filtered), 
               col=ggthemes_data$few$colors$Dark[3,2][[1]],
               alpha=0.2, size=0.2)+
  geom_boxplot(data = boxplot_overlay_dat, aes(x=sample_size_filtered, y=value_Y*100, # *100 to get percent values
                                group=sample_size_filtered), 
               col=ggthemes_data$few$colors$Dark[4,2][[1]],
               alpha=0.2, size=0.2)+
  theme_few()+
  #scale_colour_few() +
  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
  #                                         "% var(X)" = "blue", 
  #                                         "% var(Y)" = "darkblue", 
  #                                         "Number of IVs" = "black"
  #                                         ))
  #
  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                           "Number of IVs" = "black"),
                     # order
                     breaks = c( "Number of IVs", "% var(X)", "% var(Y)", "% var(U)")
  ) -> plot_pve_simulated

filename_pveplot_sim <- paste0("../output/plots/", Sys.Date(), "_figure_simplot_pve.png")
ggsave(filename = filename_pveplot_sim, plot = plot_pve_simulated, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```
